### 2022-01-18 README.Rmd
CD revision 1, updaating manuscript and supplement

---
title: "Hidden Talents in Context: Multiverse analysis repository"
output: github_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

library(tidyverse)
library(gt)
library(glue)
library(htmltools)
```

This repository contains data, code, and (some) output for a manuscript entitled "Hidden talents in context: Cognitive performance with abstract versus ecological stimuli among adversity-exposed youth" submitted to and under review at *Child Development*.

## IMPORTANT!!

The file sizes of many output files produced by scripts in this repository are very large. As a consequence, these output files are not tracked here (see .gitignore). To preserve the directory structure, files named `0-placeholder.txt` have been added to output directories. This ensures that scripts can reproduce and save output to the correct directory (without having to manually add them).

To reproduce this repository, see [How to reproduce this respository](#how-to-reproduce-this-repository) below.

## Preregistration

Details about the participants, methods, procedures, and analysis plans can be found on the Open Science Framework (OSF). Below are the relevant links for different aspects of the project:

-   [OSF project](https://osf.io/6r95z/)
-   [Original preregistration](https://osf.io/d2gz6/)
-   [Overview of measures](https://osf.io/5kawy/), which fully describes what was collected (including measures not used/analyzed in this project). Variable codebooks for the current project are contained in this repository (see Codebooks below).
-   [Final preregistration](https://osf.io/4vsnz/) which also contains a preliminary write-up of the Methods for an eventual publication
-   [Secondary analysis plan - set 1](https://osf.io/7fu35/)
-   [Secondary analysis plan - set 2](https://osf.io/wcauf/)

## Directory Structure

The names of each folder are intended to be self-explanatory. There are six top-level folders to organize the inputs and outputs of this project:

1.  `codebooks/`: lists of variable names, labels, and value labels (where applicable).
2.  `data/`: data, stored as an `.Rdata` file and `.csv` files.
3.  `manuscript/`: a manuscript written in R markdown for submission to a journal.
4.  `multiverse-objects/`: all `.Rdata` files containing objects necessary for performing multiverse analyses and storing intermediary results.
5.  `scripts/`: R-scripts that read, analyze, and produce all multiverse objects.
6.  `supplement/`: a supplemental text (to be submitted with the manuscript) documenting all secondary analyses in detail.

Below is a simple visualization of the full directory structure.

```{r}
fs::dir_tree(".", regexp = "^README|.Rproj$|log$|tex$|_files|resources", invert = T)
```

Each of the top-level folders are explained in more detail below:

## Codebooks

These are lists of variable names, labels, and value labels (where applicable). Their names correspond to the versions of the data they document (see Data below).

## Data

The raw data for this project contain all variables necessary to produce all multiverse outputs (cleaned data, multiverse datasets, effect sizes, plots). The raw data file is the only file that depends on other files/scripts to create, which saved elsewhere (for confidentiality reasons).

There are four versions of the data, which can be loaded as `R` objects or read/used as `.csv` files.

1.  The raw data
2.  The data after all *non-arbitrary* decisions were executed
3.  A version of the previous step with overall composites for the IVs
4.  A version of the data that contains variables for the multiverse analysis, only

All versions and codebooks are stored in `data/data2_analysis.Rdata`. In addition, the following `.csv` files were created to represent each version of the data separately. The prefixed numbers refer to the step in the process of going from raw to multiverse-ready data.

- `1-raw.csv`
- `2-non_arbitrary.csv`
- `3-composites.csv`
- `4-multiverse_base.csv`

A codebook for these data, including variable labels and value labels (if applicable), can be found in `codebook.csv` at the top level of this directory.

## Manuscript

This directory contains the manuscript for this project. It was written in R Markdown and is 100% reproducible. The `.Rmd` fie contains the written text with all figures, tables, and in-text statistics. All outputs reported in the manuscript were staged prior to compiling into a `.docx` document and stored in a `.Rdata` file called `staged-objects.Rdata`. The R Markdown document reads these staged files to make the overall document easier to read. The scripts that produce the staged objects are located in the `manuscript/scripts/` directory and are named to be self-explanatory. The `rmd-staging.R` script reads necessary data/objects from the `multiverse-objects/` directory and sources the other scripts to produce the objects needed for the manuscript.

## Multiverse-objects

There are four steps to doing the multiverse:

1. Specifying a grid of all possible arbitrary data processing decisions
2. Creating a "multiverse" of alternative data sets one could analyze
3. Analyzing each "universe" within the multiverse
4. Extracting the relevant information from the analyses (e.g., effect sizes, p-values, etc.)
5. Visualizing results or summarizing them in tables

This folder contains all objects necessary for representing each step. Specifically:

- `1-multiverse-datasets/`: multiverse data sets with each data universe stored in a list. There is one list used for the composites and one list used for all components of each IV.
- `2-primary-analyses/`: contain the multiverse of full results and extracted effects for the primary confirmatory and exploratory results
- `3-secondary-analyses/`: contain the same objects as above for each set of secondary analyses.

## Data Processing Scripts

There are four types of R-scripts in this repository, each with a separate folder.

- `0-functions/`: Custom R-functions written for this project
- `1-data-prep/`: Data processing scripts
- `2-primary-analyses/`: Primary analysis scripts
- `3-secondary-analyses/`: Secondary analysis scripts

Each script takes an input(s) and produces output(s). The tables below provides an overview of the inputs and outputs of each script.

```{r}
file_list <- list.files("scripts", pattern = ".R$", full.names = T, recursive = T) %>% 
  map_df(function(x){
    script_text <- 
      read_lines(x) %>% 
      str_subset("load\\(\".*Rdata\"\\)|save.image\\(.*\\)|file = .*Rdata|filename = .*pdf|write_csv.*csv") %>% 
      str_remove_all(" |\"|\\(|\\)|,") %>% 
      str_c(collapse = " ") %>% 
      str_replace_all("load","input: ") %>% 
      str_replace("write_csv|save.imagefile=|save.image|filename=|file=","~output: ") %>% 
      str_replace_all("write_csv|save.imagefile=|save.image|filename=|file=","output: ") %>% 
      str_split("~", simplify = T)

    tibble(
      dir       = x %>% str_remove_all("scripts/") %>% str_remove_all("/.*") %>% str_remove("^\\d-"),
      script    = x %>% str_remove_all(".*/"),
      raw       = script_text %>% str_squish() %>% str_c(collapse = ", "),
      n_input   = str_count(script_text[1], "input"),
      input1    = script_text[1],
      input2    = input1 %>% str_remove_all("input: ") %>% str_trim(),
      input     = input2 %>% str_split(" ", simplify = T) %>% str_remove_all(".*/") %>% str_c(collapse = "<br> ") %>% ifelse(. == "<br> ","",.),
      n_output  = str_count(script_text[2], "output") %>% ifelse(is.na(.), 0, .),
      output1   = script_text[2],
      output2   = output1 %>% str_remove_all("output: ") %>% str_trim(),
      output    = output2 %>% str_split(" ", simplify = T) %>% str_remove_all(".*/") %>% str_c(collapse = "<br> ") %>%  ifelse(. == "<br> NA","",.)
    )
  }) 
```

### Data Prep

```{r, results='asis'}
data_prep <- 
  file_list %>% 
  select(dir, script, input, output) %>% 
  filter(dir == "data-prep") %>% 
  select(-dir)

cat(
  "|script|input|output|",
  "|------|-----|------|",
  glue::glue_data(
    data_prep,
    "|{script}|{input}|{output}|"
  ),
  sep = "\n"
)
```

### Primary Analyses

```{r, results='asis'}
primary_analyses <- 
  file_list %>% 
  select(dir, script, input, output) %>% 
  filter(dir == "primary-analyses") %>% 
  select(-dir)

cat(
  "|script|input|output|",
  "|------|-----|------|",
  glue::glue_data(
    primary_analyses,
    "|{script}|{input}|{output}|"
  ),
  sep = "\n"
)
```

### Secondary Analyses

```{r, results='asis'}
secondary_analyses <- 
  file_list %>% 
  select(dir, script, input, output) %>% 
  filter(dir == "secondary-analyses") %>% 
  select(-dir)

cat(
  "|script|input|output|",
  "|------|-----|------|",
  glue::glue_data(
    secondary_analyses,
    "|{script}|{input}|{output}|"
  ),
  sep = "\n"
)
```

## Supplement

This directory contains the supplement to the manuscript described above. The supplement was produced using the same approach as the manuscript: it was written in R Markdown, is 100% reproducible, and all outputs were staged prior to compiling into a `.docx`. The staged objects are stored in `supplement/staged-objects.Rdata`. The scripts that produce the staged objects are located in the `supplement/scripts/` directory and are named to be self-explanatory. The `rmd-staging.R` script reads necessary data/objects from the `multiverse-objects/` directory and sources the other scripts to produce the objects needed for the supplement.

## How to reproduce this repository

To reproduce this repository, download all files locally. You can do this by cloning the repository or downloading the files directly as a .zip file. Because many of the objects are very large (i.e., those containing multiverse data and results), they are not included in this repository. However, the scripts will create all necessary outputs to reproduce all data and files. Below is the exact order these scripts should be ran:

### 1.  Data Preparation

1. `1-data-prep/1-data-non-arbitrary.R`
2. `1-data-prep/2-primary-multiverse-datasets.R`
3. `1-data-prep/3-secondary-multiverse-datasets.R`

### 2.  Primary Analyses

1. `2-primary-analyses/1-confirmatory.R`: conducts unpredictability and violence multiverse analyses for attention-shifting and working memory updating.
2. `2-primary-analyses/2-exploratory.R`: conducts poverty multiverse analyses for attention-shifting and working memory updating.
3. `2-primary-analyses/3-extract-effects.R`: extracts and compiles multiverse data for plotting and reporting.
4. `2-primary-analyses/4-bootstrap-setup.R`: sets up bootstrap samples for bootstrapping technique .
5. `2-primary-analyses/5-bootstrap-spec-analysis.R`: runs specification analyses over all bootstrapped samples to obtain overall *p*-values.

### 3. Secondary Analyses

1. Set 1
    1. `3-secondary-analyses/set-1/a1-analysis-ses-vio.R`: conducts secondary analysis labeled "Comparing Violence and Poverty Exposure Interactions with Task-Version" in the manuscript
    2. `3-secondary-analyses/set-1/a2-extract-effects.R`: extracts and compiles multiverse data for plotting and reporting for the analysis labeled "Comparing Violence and Poverty Exposure Interactions with Task-Version" in the manuscript.
    3. `3-secondary-analyses/set-1/b1-analysis-ses-components.R`: conducts an analysis of the components of poverty, which is redundant with the secondary analysis labeled "Exploring Components of Adversity Dimensions", which is conducted below in Set 2. 
    4. `3-secondary-analyses/set-1/b2-extract-effects.R`: extracts effects for poverty component multiverse analyses.
2. Set 2
    1. `3-secondary-analyses/set-2/a1-analysis-adversity-components.R`: conducts an analysis of the components of all adversity composites for the secondary analysis labeled "Exploring Components of Adversity Dimensions" (this is an expansion of the above secondary analysis for the components of poverty in Set 1) 
    2. `3-secondary-analyses/set-2/a2-extract-effects.R`: extracts effects for adversity component multiverse analyses.

### 4. Manuscript

1. `manuscript/scripts/rmd-staging.R`: loads all necessary data and multiverse objects, computes all necessary in-text statistics, and sources the following scripts to create tables and figures for the manuscript stored in `manuscirpt/staged-objects.Rdata`:
    - `table1.R`
    - `table2.R`
    - `table3.R`
    - `figure2-3.R`
2. `manuscript/manuscript.Rmd`: loads `manuscript/staged-objects.Rdata` and knits together the written manuscript and all statistics, tables, and figures computed by the staging script above.

### 5. Supplement

1. `supplement/scripts/rmd-staging.R`: loads all necessary data and multiverse objects and sources the following scripts to create tables and figures for the supplement stored in `supplement/staged-objects.Rdata`:
    - `1-covariates.R`
    - `2-secondary1.R`
    - `3-secondary2.R`
    - `4-tables.R`
    - `5-tables-r1.R`
2. `supplement/supplement.Rmd`: loads `supplement/staged-objects.Rdata` and knits together the written supplement and all tables and figures computed by the staging script above.