### 2022-03-16 manuscript/tables/table-1.Rmd
make separate .docx files for tables

---
output:
  officedown::rdocx_document:
    reference_docx: ../reference-doc.docx
    page_margins:
      bottom: 1
      top: 1
      right: 1
      left: 1
      header: .5
      footer: 0
      gutter: 0
    plots:
      align: center
      caption:
        style: Image Caption
        pre: 'Figure '
        sep: '. '
    tables:
      caption:
        style: Table Caption
        pre: 'Table '
        sep: '. '
---

```{r setup, include=FALSE}
# set up chunk options
knitr::opts_chunk$set(
  echo = F,
  fig.align = "center",
  fig.pos = "!t", 
  out.extra = "",
  fig.show = "asis",
  message = FALSE,
  tab.topcaption = T,
  warning = FALSE
)

# load libraries
library(tidyverse)
library(officer)
library(flextable)

# load all data and objects
load("../staged-objects.Rdata")

# set up flextable for tables
set_flextable_defaults(
  font.family = "Times", 
  font.size = 10,
  font.color = "black",
  line_spacing = 1,
  padding.bottom = 1, 
  padding.top = 1,
  padding.left = 1,
  padding.right = 1
)
```

```{r table1, tab.id = "table1", tab.cap.style = "Table Caption", tab.cap="Bivariate correlations and descriptive statistics for adversity measures."}
table.01.0 %>% 
  add_column(.after = 4,"empty1" = "") %>% 
  add_column(.after = 8,"empty2" = "") %>% 
  flextable() %>%
  width("empty1", width = .2) %>% 
  width("empty2", width = .2) %>% 
  width(j = 1, width = .75) %>% 
  set_header_labels(Variable = "", empty1 = "", empty2 = "") %>% 
  add_header_row(
    values = c(" ", "Unpredictability", " ", "Violence", " ", "Poverty"),
    colwidths = c(1, 3, 1, 3, 1, 4)
  ) %>% 
  border_remove() %>% 
  border(i = 1, j = c("1","2","3","4","5","6","7","8","9","10"), border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 1, border.top = fp_border_default(), part = "header") %>% 
  border(i = 2, border.bottom = fp_border_default(), part = "header") %>% 
  border(i = 16, border.bottom = fp_border_default(), part = "body") %>% 
  align(i = 1:2, align = "center", part = "header") %>% 
  bold(i = 1:2, part = "header") %>% 
  set_table_properties(width = 1, layout = "autofit") %>% 
  add_footer_row(
    values = " ",
    colwidths = 13
  ) %>% 
  add_footer_row(
    values = " ",
    colwidths = 13
  ) %>% 
  compose(
    i = 1, j = 1, 
    as_paragraph(as_i("Note: "), "Column headers denote adversity constructs and rows indicate components. The upper triangle contains sample sizes for each correlation."), 
    part = "footer"
  ) %>% 
  compose(
    i = 2, j = 1, 
    as_paragraph(as_i("* p "), "< .05, ", as_i("** p "), "< .01"), 
    part = "footer"
  ) 
```